---
title: "FMarker-list curation (GBS) step 1"
author: "Luc Baudouin"
date: "15 février 2016"
output: pdf_document
---

First step: Tassel was executed and produced a VCF file. $\chi^2$ scores and P_values were computed at each locus. These values along with marker information and genotypic data were included into a new file. This step can take one or more input files (eg. because data were treated in parallel or to save memory space). For each input file, a genotypic file and a marker info file are created.

# Preliminary





```{r utilities,echo=FALSE}
topath<-function(path,file) paste(path, file,sep="/")
# Note: Set the working directory as the one where the present file is located. 
here<-getwd()
source(topath(here,"tassel_to_J_parameters.R"))
if(!dir.exists(where_out_1)) dir.create(where_out_1)
```

# Information on the dataset

`r comment0`

# Additional comments 

`r comment1`

# procedure

```{r functions, echo=FALSE}
# The functions used here are in a special directory
function_file<-list.files(where_functions)
function_file<-function_file[grepl(".[rR]$",function_file,perl=TRUE)]

bid<-sapply(paste(where_functions,"/",function_file, sep=""),source,echo=FALSE,verbose=FALSE)

```

## execution

The first test consisted in retaining only those markers whose distribution is not too far from 50-50. 

```{r data_entry, echo=FALSE}
# convert data files and save results
check<-convert_data(file.value, file.in, file.ident, where_in, P.value.min, where_out_1)
message<-ifelse(length(unique(check[,"individuals"]))!=1,"problem with number of individuals","ok")

initial_mark<-sum(check[,"before_mark"])
final_mark<-sum(check[,"after_mark"])
initial_scaff<-sum(check[,"before_scaff"])
final_scaff<-sum(check[,"after_scaff"])
ratio_mark<-final_mark/initial_mark
ratio_scaff<-final_scaff/initial_scaff
```


This step ended with the status `r message`.
Data with p.values below `r P.value.min` were filtered out. A negative value means no filtering. 

* The number of markers was reduced  from `r formatC(initial_mark,format="f",big.mark = ",",digits=0)` to `r formatC(final_mark,format="f",big.mark = ",",digits=0)`. `r round(ratio_mark*100,2)` % of loci were conserved.
* The number of scaffolds reduced from `r formatC(initial_scaff,format="f",big.mark = ",",digits=0)` to `r formatC(final_scaff,format="f",big.mark = ",",digits=0)`. `r round(ratio_scaff*100,2)` % of scaffolds were conserved.


Data were in file(s)    
`r file.in `

in folder    
`r where_in`

Created files     

`r paste(c(file.value,file.ident), collapse="\n\n")`

in folder    

`r where_out_1`
