---
title: "Final data output (GBS)"
author: "Luc Baudouin"
date: "18 mai 2016"
output: pdf_document
---

# Introduction
Step 5: This is the last step. It creates a data file that can be used by Joinmap .


# Preliminary

```{r get_parameters,echo=FALSE}
topath<-function(path,file) paste(path, file,sep="/")
# Note: Set the working directory as the one where the present file is located. 
here<-getwd()
source(topath(here,"tassel_to_J_parameters.R"))
if(!dir.exists(where_out_5)) dir.create(where_out_5)

#require("xtable",quietly=TRUE)
```

# Information on the dataset

`r comment5`

Data are coded as "`r genotype_code`"

## data files

Genotyping data are in file "`r reduced_data`",  directory "`r where_out_4 `".

Identifiers are in file "`r file.ident.3`",  directory "`r where_out_4 `".

Locus column is `r marker_column`, scaffold column is `r scaffold_column`, position column is `r position_column`.

Output directory is `r where_out_4`.

## Source files
```{r functions,echo=FALSE}

# The functions used here are in a special directory
function_file<-list.files(where_functions)
function_file<-function_file[grepl(".[rR]$",function_file,perl=TRUE)]

bid<-sapply(paste(where_functions,"/",function_file, sep=""),source,echo=FALSE,verbose=FALSE)
full_path<- function(place,file) paste(place,file,sep="/")
output_data<-function(Table, place, file) write.table(cbind(index=1:nrow(Table),loc=rownames(Table),Table),full_path(place,file),sep="\t",row.names=FALSE,quote=FALSE)
code_U <- strsplit(genotype_code,split = "")[[1]]
code_L <- tolower(code_U)
val1 <- sort(c(code_L, code_U))
  unknown <- c(code_L[4],code_U[4])
```

The following function files were loaded:

```{r load_function,echo=FALSE}
noquote(function_file)
```

# procedure

## Opening the data files

```{r, open_files,echo=FALSE}
# Opening the genotypic data file
reduced_data<-full_path(where_out_4,reduced_data)
dat<-read.table(reduced_data,sep="\t",stringsAsFactors = FALSE,header=TRUE)
dat<-dat[,-1]#une colonne pour rien A corriger ? l'avenir
reduced_info<-full_path(where_out_4,reduced_info)
info<-read.table(reduced_info,sep="\t",stringsAsFactors = FALSE,header=TRUE)

indiv_select<-full_path(where_out_4,indiv_select)
rem<-read.table(indiv_select,sep="\t",stringsAsFactors = FALSE,header=TRUE)

#Locus name
loc<-dat[,1]
rownames(info)<-info$locus
genot<-dat[,-1]
rownames(genot)<-loc
if(removed>0) to_remove<-as.character(rem$remove[1:removed]) else to_remove<-NULL

keep_ind<-setdiff(names(genot),to_remove)
 zloc <-
    t(apply(genot[,keep_ind],1,function(y)
      apply(sapply(val1, function(x)
        match(y, x,nomatch = 0)),2,sum)))

  cmiss.loc <- apply(zloc[,unknown],1,sum)
 
  pmiss.loc <- cmiss.loc / length(keep_ind)
  p.keep.loc <- pmiss.loc[pmiss.loc < max_miss_loc]
  
  out<-tolower(sapply(genot[names(p.keep.loc),keep_ind], as.character))
  rownames(out)<-names(p.keep.loc)
  output_data(out,where_out_5,genot_file_5)
  output_data(info[names(p.keep.loc),],where_out_5,infoloc_5)

```
Data were in file(s)

`r reduced_data`

`r reduced_info`

in folder    

`r where_out_4`
 
 
Created files
 
`r paste(file.i, collapse="\n\n")`
 
in directory    

`r where_out_3`


