---
title: "resultats tassel part 2"
author: "Luc Baudouin"
date: "15 février 2016"
output: pdf_document
---


Second step: At this stage, the genotypes are assigned to the parents. Only loci whose segregation is compatible with the mating plan are retained. Statistics are computed on the disjunction and the results are written into the locus info file. The columns of the genotype table are reordered. First the parents, then the descendents in order of increasing number of missing data. All genotypeing files and all locus information file are merged at this step.  

# Preliminary


```{r get_parameters,echo=FALSE}
topath<-function(path,file) paste(path, file,sep="/")
# Note: Set the working directory as the one where the present file is located. 
here<-getwd()
source(topath(here,"tassel_to_J_parameters.R"))
if(!dir.exists(where_out_2)) dir.create(where_out_2)


```

# Information on the dataset

`r comment0`

# Additional comments

`r comment2`

# procedure



## opening the data file

```{r data_entry,echo=FALSE}

#Remettre l'ouverture des fichiers. Fusionner ?
n_value<-length(file.value)
n_id<-length(file.ident)
if(n_id!=n_value) stop("Numbers of marker identity and genotype files differ")
value<-NULL
ident<-NULL
for(i in 1:n_value){ 
  value<-rbind(value, read.table(topath(where_out_1,file.value[i]),header = TRUE,stringsAsFactors = FALSE))
  ident<-rbind(ident, read.table(topath(where_out_1,file.ident[i]),header = TRUE,stringsAsFactors = FALSE))
} 

  markers<-rownames(value) 
  indiv<-colnames(value)
 
# dim(value)
# length(markers)

progeny<-indiv[!(indiv %in% parent)&!(indiv %in% others)]


```

```{r parent_genotype,echo=FALSE}
#parent
#Valid: the parents have expected genotypes.
if(has.hetero){ # heterozygote present 
  if(has.homo){ # both parents
    expected<-cbind(c("A/C","A/G","A/T","C/A","G/A","T/A",
                      "A/C","C/A","C/G","C/T","G/C","T/C",
                      "A/G","C/G","G/A","G/C","G/T","T/G",
                      "A/T","C/T","G/T","T/A","T/C","T/G"),
                    c("A/A","A/A","A/A","A/A","A/A","A/A",
                      "C/C","C/C","C/C","C/C","C/C","C/C",
                      "G/G","G/G","G/G","G/G","G/G","G/G",
                      "T/T","T/T","T/T","T/T","T/T","T/T"))
    colnames(expected)<-parent
    expect<-apply(expected,1,paste,collapse=" ")
    test<-apply(value[,parent],1,paste,collapse=" ")
    tp1<-table(value[,parent[1]])
    tp2<-table(value[,parent[2]])
    glist<-sort(union(names(tp1),names(tp2)))
    out1<-matrix(0, 2,length(glist),dimnames=list(parent,glist))
    out1[parent[1],names(tp1)]<-tp1
    out1[parent[2],names(tp2)]<-tp2

    rownames(out1)<-parent
    capt<-"Frequency of genotypes (%) in both parents"
  }else{ # only heterozygote
    expect<-expected<-c("A/C","A/G","A/T","C/A","C/G","C/T",
                     "G/A","G/C","G/T","T/A","T/C","T/G")
    test<-value[,parent[1]]
    out1<- t(table( value[,parent[1]]))
    rownames(out1)<-parent[1]
    capt<-"Frequency of genotypes (%) in heterozygous parent"
  } # end homo 1
}else{ # no heeterozygote
  if(has.homo){#only homozygote
    expect<-expected<-c("A/A","C/C","G/G","T/T")
    test<-value[,parent[2]]
    out1<- t(table( value[,parent[2]]))
    rownames(out1)<-parent[2]
    capt<-"Frequency of genotypes (%) in homozygous parent"
  }else{ # no parent available
    stop("at least one parent should be genotyped")
  }  #end homo 2
  
} # end hetero


```

```{r test,echo=FALSE}

valid<-test %in% expect

loc_init<-nrow(value)
scaff_init<-length(unique(as.character(ident[[scaffold_column]]))) 

value<-value[valid,] 
ident<-ident[valid,] 
loc_final<-nrow(value) 
ratio_loc<-loc_final/loc_init
scaff_final<-length(unique(as.character(ident[[scaffold_column]])))
ratio_scaff<-scaff_final/scaff_init
ident$Hyb<-value[[parent[1]]] 
ident$Dwarf<-value[[parent[2]]] 
#write.table(cbind(marker=rownames(value),value),topath(where_out,file.value.2),sep="\t",row.names=FALSE,quote=FALSE)


```

```{r last_step,echo=FALSE}

test.indf <- function(ind,genot) {
  a <- as.character(genot[,ind])
  l <- length(a)
  miss <- sum(a == "./.")
  miss / l
}
  
test.par <- sapply(parent, function(x)
  test.indf(x, value))
test.ind <- sapply(progeny, function(x)
  test.indf(x, value))
sti <- 1 - sort(test.ind,decreasing = FALSE)
sti2 <- sti[sti > cut.ind]
retain <- names(sti2)
remove<-setdiff(progeny,retain)
if (draw_plot) {
  plot(
    1:length(progeny),sti,type = "l",ylim = c(0,1),xlab = "individuals",ylab =
      "valid data"
  )
  abline(h = cut.ind, col = "blue")
  abline(v = length(retain), col = "blue")
}
#length(retain)
value2 <- cbind(value[,parent],value[,retain])
write.table(
  cbind(rownames(value2),value2),topath(where_out_2,file.value.2),sep = "\t",row.names =
    FALSE,quote = FALSE
)
test.dist <- function(mark, genot, ident,choice = retain) {
  a <- as.character(genot[mark,choice])
  hyb <- sum(a == as.character(ident[mark,"Hyb"]))
  dwarf <- sum(a == as.character(ident[mark,"Dwarf"]))
  miss <- sum(a == "./.")
  c(
    D = dwarf, H = hyb, M = miss, wrong = length(a) - hyb - miss - dwarf
  )
  
}
test1 <-
  t(sapply(1:nrow(value), function(x)
    test.dist(x, value2,ident)))
ident <- cbind(ident,test1)# dim(ident)
ident$total <- apply(test1,1,sum)
ident$good <- apply(test1[,1:2],1,sum)
ident$f_good <- ident$good / ident$total
ident$is_good <- ident$f_good > few_good_data
ident$f_D<-ident$D/ident$good
ident$balanced <-
  ident$f_D > balanced & ident$f_D < 1 - balanced
ident$third_allele <- ident$wrong > many_wrong
write.table(
  ident,topath(where_out_2,file.ident.2),sep = "\t",row.names = FALSE,quote =
    FALSE
)
  
```
End of step 2
 
* The number of markers was reduced  from `r formatC(loc_init,format="f",big.mark = ",",digits=0)` to `r formatC(loc_final,format="f",big.mark = ",",digits=0)`. `r round(ratio_loc*100,2)` % of loci were conserved.
* The number of scaffolds reduced from `r formatC(scaff_init,format="f",big.mark = ",",digits=0)` to `r formatC(scaff_final,format="f",big.mark = ",",digits=0)`. `r round(ratio_scaff*100,2)` % of scaffolds were conserved.
* `r length(remove)` individuals were removed because they don't reach `r cut.ind*100` % valid data:  `r remove`. The number of remaining progenies is `r length(retain)`.


Data were in file(s)

`r file.value `

`r file.ident `

in folder    

`r where_out_1`

Created files

`r paste(dir(where_out_2), collapse="\n\n")`

in directory

`r where_out_2`

 
