---
title: "Marker-list curation (GBS) step 3"
author: "Luc Baudouin"
date: "15 février 2016"
output: pdf_document
---
Codes the genotype file in a form usable by mapping software. Parents are removed from the output file.

# Preliminary

```{r get_parameters,echo=FALSE}
topath<-function(path,file) paste(path, file,sep="/")
# Note: Set the working directory as the one where the present file is located. 
here<-getwd()
source(topath(here,"tassel_to_J_parameters.R"))
if(!dir.exists(where_out_3)) dir.create(where_out_3)

suppressWarnings(require("xtable",quietly=TRUE))
```

# Information on the dataset

`r comment0`

# Additional comments

`r comment3`

# procedure



```{r utilities,echo=FALSE}
code <- function(val,id,parent,gcode="ABHU") {
  code_l<-strsplit(gcode,split =  "")[[1]]
  val2 <- rep(code_l[4],length(val))
  names(val2)<-names(val)
  val2[val == id$Hyb] <- code_l[3]
  val2[val == id$Dwarf] <- code_l[1]
  val2[!(names(val2) %in% parent)][-1]
}


```


## opening the data file

```{r data_entry,echo=FALSE}
#Remettre l'ouverture des fichiers. Fusionner ?
n_value<-length(file.value)
n_id<-length(file.ident)
if(n_id!=n_value) stop("Numbers of marker identity and genotype files differ")
value<-NULL
ident<-NULL
for(i in 1:n_value){ 
  value<-rbind(value, read.table(topath(where_out_2,file.value.2[i]),header = TRUE,stringsAsFactors = FALSE))
  ident<-rbind(ident, read.table(topath(where_out_2,file.ident.2[i]),header = TRUE,stringsAsFactors = FALSE))
} 

  markers<-value[[1]] 
  indiv<-colnames(value) 
 
 

progeny<-indiv[!(indiv %in% parent)&!(indiv %in% others)] 


```

Sample data

```{r recode,echo=FALSE}

keep=ident$is_good&ident$balanced
loc_init<-length(keep)
loc_final<-sum(keep)
scaff_init<-length(unique(ident[[scaffold_column]]))
scaff_final<-length(unique(ident[[scaffold_column]][keep]))
ratio_loc<-loc_final/loc_init
ratio_scaff<-scaff_final/scaff_init

coded<-t(sapply(1:length(markers), function(x) code(value[x,],ident[x,],parent)) )
write.table(cbind(marker=markers[keep],coded[keep,]),topath(where_out_3,file.value.3),sep="\t",row.names=FALSE,quote=FALSE)
write.table(ident[keep,],topath(where_out_3,file.ident.3),sep="\t",row.names=FALSE,quote=FALSE)
#rownames(coded)<-markers
noquote(cbind(markers[1:10],coded[1:10,1:10]))
```

End of step 3
 
* The number of markers was reduced  from `r formatC(loc_init,format="f",big.mark = ",",digits=0)` to `r formatC(loc_final,format="f",big.mark = ",",digits=0)`. `r round(ratio_loc*100,2)` % of loci were conserved.
* The number of scaffolds reduced from `r formatC(scaff_init,format="f",big.mark = ",",digits=0)` to `r formatC(scaff_final,format="f",big.mark = ",",digits=0)`. `r round(ratio_scaff*100,2)` % of scaffolds were conserved.

Data were in file(s)

`r file.value.2 `

`r file.ident.2 `

in folder    

`r where_out_2`
 
 
Created files
 

`r file.value.3 `

`r file.ident.3 `
 
in folder    

`r where_out_3`

